---
title: "RWorksheet_Camarista#4c"
author: "John Lyxton Camarista"
date: "2024-11-01"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##1. Use the dataset mpg
  - a. Show your solutions on how to import a csv file into the environment.
```{r}
library(readr)


mpg_data <- read_csv("E:/Github/Data Science Worksheets/DataScience_Worksheets_Camarista/Worksheet#4/mpg.csv")  

head(mpg_data)

```
  - b. Which variables from mpg dataset are categorical?  
    * The categorical variables are: manufacturer, model, year, trans, drv, fl, and class.    
  - c. Which are continuous variables?    
    *The continous variables are: displ, cyl, cty, and hwy.
    
##2. Which manufacturer has the most models in this data set? Which model has the most variations? Show your answer.
  - a. Group the manufacturers and find the unique models. Show your codes and result.   
```{r}
# Load necessary library
library(dplyr)

# Get unique models for each manufacturer
unique_models <- mpg_data %>%
  select(manufacturer, model) %>%
  distinct() %>%
  arrange(manufacturer)

# Display the result
unique_models

```
  - b. Graph the result by using plot() and ggplot(). Write the codes and its result.   
```{r}
library(dplyr)

model_count <- mpg_data %>%
  select(manufacturer, model) %>%
  distinct() %>%
  group_by(manufacturer) %>%
  summarise(model = n())

model_count$manufacturer <- factor(model_count$manufacturer, levels = unique(model_count$manufacturer))

plot(
  model_count$manufacturer,
  model_count$model,
  type = "h",  lines,
  main = "Number of Unique Models by Manufacturer",
  xlab = "Manufacturer",
  ylab = "Number of Unique Models",
  col = "skyblue",
  las = 2,  
  lwd = 10    
)
```
```{r}
library(ggplot2)

ggplot(model_count, aes(x = reorder(manufacturer, model), y = model)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = "Number of Unique Models by Manufacturer",
       x = "Manufacturer",
       y = "Number of Unique Models") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


##3. Same dataset will be used. You are going to show the relationship of the modeland the manufacturer.

##4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result

##5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using geom_point with aesthetic color = engine displacement. Title should be “Relationship between No. of Cylinders and Engine Displacement”.

##6. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped it with a continuous variable you have identified in #1-c. What is its result? Why it produced such output?

##7. Import the traffic.csv onto your R environment.

##8. From alexa_file.xlsx, import it to your environment