---
title: "RWorksheet_Camarista#4c"
author: "John Lyxton Camarista"
date: "2024-11-01"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##1. Use the dataset mpg
  - a. Show your solutions on how to import a csv file into the environment.
```{r}
library(readr)


mpg_data <- read_csv("E:/Github/Data Science Worksheets/DataScience_Worksheets_Camarista/Worksheet#4/mpg.csv")  

head(mpg_data)

```
  - b. Which variables from mpg dataset are categorical?  
    * The categorical variables are: manufacturer, model, year, trans, drv, fl, and class.    
  - c. Which are continuous variables?    
    *The continous variables are: displ, cyl, cty, and hwy.
    
##2. a. Which manufacturer has the most models in this data set? Which model has the most variations? Show your answer.
  - a. Group the manufacturers and find the unique models. Show your codes and result.   
```{r}
# Load necessary library
library(dplyr)

# Get unique models for each manufacturer
unique_models <- mpg_data %>%
  select(manufacturer, model) %>%
  distinct() %>%
  arrange(manufacturer)

# Display the result
unique_models

```
  - b. Graph the result by using plot() and ggplot(). Write the codes and its result.   
```{r}
library(dplyr)

model_count <- mpg_data %>%
  select(manufacturer, model) %>%
  distinct() %>%
  group_by(manufacturer) %>%
  summarise(model = n())

model_count$manufacturer <- factor(model_count$manufacturer, levels = unique(model_count$manufacturer))

plot(
  model_count$manufacturer,
  model_count$model,
  type = "h",  lines,
  main = "Number of Unique Models by Manufacturer",
  xlab = "Manufacturer",
  ylab = "Number of Unique Models",
  col = "skyblue",
  las = 2,  
  lwd = 10    
)
```
```{r}
library(ggplot2)

ggplot(model_count, aes(x = reorder(manufacturer, model), y = model)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = "Number of Unique Models by Manufacturer",
       x = "Manufacturer",
       y = "Number of Unique Models") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


##2. b. Same dataset will be used. You are going to show the relationship of the modeland the manufacturer.
  - a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?    
    * Each point represents an individual observation in the mpg dataset, with each combination of model and manufacturer plotted as a point.
```{r}
ggplot(mpg, aes(model, manufacturer)) + geom_point()
```
  - b. For you, is it useful? If not, how could you modify the data to make it more informative?    
  * Is not particularly useful in its current form because it’s cluttered due to the large number of unique models on the x-axis.   
  * A bar chart or heatmap showing counts or averages would provide a clearer, more informative view than the original scatter plot.
  
##3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and its results.
```{r}
ggplot(head(mpg_data, 20), aes(x = model, y = year)) +
  geom_point() +
  labs(title = "Model vs Year (Top 20 Observations)", x = "Model", y = "Year") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
  
##4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result.    
  - a. Plot using geom_bar() using the top 20 observations only. The graphs should have a title, labels and colors. Show code and results.
```{r}
library(dplyr)
library(ggplot2)

model_counts <- mpg %>%
  group_by(model) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

top_20_models <- model_counts %>%
  head(20)

ggplot(top_20_models, aes(x = reorder(model, count), y = count, fill = model)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Top 20 Car Models by Count",
    x = "Model",
    y = "Number of Cars"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_viridis_d()

```


##5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using geom_point with aesthetic color = engine displacement. Title should be “Relationship between No. of Cylinders and Engine Displacement”.   
  - a. How would you describe its relationship? Show the codes and its result.
```{r}
ggplot(mpg_data, aes(x = cyl, y = displ, color = displ)) +
  geom_point() +
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders", y = "Engine Displacement")

```

##6. a. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped it with a continuous variable you have identified in #1-c. What is its result? Why it produced such output?
```{r}
ggplot(mpg_data, aes(x = displ, y = hwy, color = cty)) +
  geom_point() +
  labs(title = "Engine Displacement vs Highway MPG",
       x = "Engine Displacement", y = "Highway MPG", color = "City MPG")

```

##6. Import the traffic.csv onto your R environment.
```{r}
library(readr)

traffic_data <- read_csv("E:/Github/Data Science Worksheets/DataScience_Worksheets_Camarista/Worksheet#4/traffic.csv")

traffic_data$DateTime <- as.POSIXct(traffic_data$DateTime, format = "%d/%m/%Y %H:%M", tz = "UTC")
```
  - a. How many numbers of observation does it have? What are the variables of the traffic dataset the Show
your answer.
```{r}
traffic_obs <- nrow(traffic_data)
print(paste("There are", traffic_obs, "observations"))
print("The variables in the traffic dataset are: ")
names(traffic_data)
```
  - b. subset the traffic dataset into junctions. What is the R codes and its output?
```{r}
library(dplyr)

# Create subsets for each junction
junction_data <- traffic_data %>%
  group_by(Junction) %>%
  group_split()
print(junction_data[[1]])
```
  - c. Plot each junction in a using geom_line(). Show your solution and output.
```{r}
library(ggplot2)
  ggplot(traffic_data, aes(x = DateTime, y = Vehicles, color = factor(Junction))) +
    geom_line() +
    labs(title = "Traffic Counts by Junction Over Time",
         x = "Date and Time",
         y = "Number of Vehicles",
         color = "Junction") +
    theme_minimal()

```

##7. From alexa_file.xlsx, import it to your environment
  - a. How many observations does alexa_file has? What about the number of columns? Show your solution
and answer.   
```{r}
library(readxl)

alexa_data <- read_excel("E:/Github/Data Science Worksheets/DataScience_Worksheets_Camarista/Worksheet#4/alexa_file.xlsx")

alexa_obs <- nrow(alexa_data)
alexa_num_cols <- ncol(alexa_data)

print(paste("There are ", alexa_obs, "observations in the data"))
print(paste("There are ", alexa_num_cols, "columns int the data"))
```
  - b. group the variations and get the total of each variations. Use dplyr package. Show solution and answer.
```{r}
library(dplyr)

alexa_variation_counts <- alexa_data %>%
  group_by(variation) %>%
  summarise(Total = n())

print(alexa_variation_counts)
```
  - c. Plot the variations using the ggplot() function. What did you observe? Complete the details of the
graph. Show solution and answer.    
```{r}
library(ggplot2)
library(RColorBrewer)

custom_palette <- colorRampPalette(brewer.pal(9, "Set1"))(20)

ggplot(alexa_variation_counts, aes(x = reorder(variation, -Total), y = Total, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Distribution of Alexa Variations",
    x = "Variation",
    y = "Total Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = custom_palette)

```
  
  - d. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the graphs. Show your answer and solution.
```{r}
alexa_data$date <- as.Date(alexa_data$date)

daily_reviews <- alexa_data %>%
  group_by(date) %>%
  summarise(review_count = n(), .groups = 'drop')

ggplot(daily_reviews, aes(x = date, y = review_count)) +
  geom_line(color = "blue") +
  labs(
    title = "Number of Verified Reviews Over Time",
    x = "Date",
    y = "Review Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

  - e. Get the relationship of variations and ratings. Which variations got the most highest in rating? Plot a graph to show its relationship. Show your solution and answer. 
```{r}
# Calculate the average rating by variation
variation_ratings <- alexa_data %>%
  group_by(variation) %>%
  summarise(Average_Rating = mean(rating, na.rm = TRUE), .groups = 'drop') %>%
  arrange(desc(Average_Rating))

# Plot the average ratings by variation
custom_palette <- colorRampPalette(brewer.pal(9, "Set1"))(20)

ggplot(variation_ratings, aes(x = reorder(variation, Average_Rating), y = Average_Rating, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average Ratings by Variation",
    x = "Variation",
    y = "Average Rating"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = custom_palette)

```
