---
title: "RWorksheet_Camarista#1"
author: "John Lyxton L. Camarista"
date: "2024-09-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Set up a vector named age, consisting of 34, 28, 22, 36, 27, 18, 52, 39, 42, 29,
35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41,
51, 35, 24, 33, 41.
a. How many data points?

    There are 34 data points.

b. Write the R code and its output.
```{r}
Age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37,
         34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41, 51, 35, 24, 33, 41)
```


```{r}
length(Age)
```

2. Find the reciprocal of the values for age.
Write the R code and its output.
```{r}
library(MASS)
reciprocal <- 1/Age
fractions(reciprocal)

```




3. Assign also new_age <- c(age, 0, age).
```{r}
new_age <- c(20, 0, 54)
```

What happen to the new_age?
    
    The new_age _____ 




4. Sort the values for age.
Write the R code and its output.

```{r}
sort(Age)
```




5. Find the minimum and maximum value for age.
Write the R code and its output.
```{r}
min(Age)
max(Age)
```




6. Set up a vector named data, consisting of 2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5,
2.3, 2.5, 2.3, 2.4, and 2.7.
a. How many data points?
    
    There are 12 data points.

b. Write the R code and its output.
```{r}
data <- c(2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5,
2.3, 2.5, 2.3, 2.4, 2.7)
length(data)
```
7. Generates a new vector for data where you double every value of the data. | 
```{r}
ageDoubled <- Age * 2
```

What happen to the data?

    The value of (Age) become doubled in the new vector (ageDoubled)
    
8. Generate a sequence for the following scenario:
8.1 Integers from 1 to 100.
```{r}
seq_1_to_100 <- 1:100
print(seq_1_to_100)
```


8.2 Numbers from 20 to 60
```{r}
seq_20_to_60 <- 20:60
print(seq_20_to_60)
```


*8.3 Mean of numbers from 20 to 60
```{r}
mean(seq_20_to_60)
```


*8.4 Sum of numbers from 51 to 91
```{r}
seq_51_to_91 <- 51:91
sum(seq_51_to_91)
```


*8.5 Integers from 1 to 1,000
a. How many data points from 8.1 to 8.4?
    8.1 = 100
    8.2 = 41
    8.3 = 41
    8.4 = 41

b. Write the R code and its output from 8.1 to 8.4.
```{r}
length(seq_1_to_100)
length(seq_20_to_60)
length(seq_20_to_60)
length(seq_51_to_91)
sum(100, 41, 41, 41)
```


c. For 8.5 find only maximum data points until 10.
```{r}
seq_1_to_1000 <- 1:1000
max_to_10 <- max(seq_1_to_1000[seq_1_to_1000 <- 10])
print(max_to_10)
```
9. *Print a vector with the integers between 1 and 100 that are not divisible by 3, 5 and
7 using filter option.
Filter(function(i) { all(i %% c(3,5,7) != 0) }, seq(100))
Write the R code and its output.
```{r}
answer <- Filter(function(i) {
  all(i %% c(3, 5, 7) != 0)
}, seq_1_to_100)
print(answer)
```
10. Generate a sequence backwards of the integers from 1 to 100.
Write the R code and its output.)
```{r}
reverse_1_to_100 <- 100:1
print(reverse_1_to_100)
```
11. List all the natural numbers below 25 that are multiples of 3 or 5.
Find the sum of these multiples.
```{r}
seq_1_to_25 <- 1:25
Filter(function(i){
  all(i %% c(3, 5) == 0)
}, seq_1_to_25)
```
a. How many data points from 10 to 11?
    -there are 125 data points from numbers 10 to 11

b. Write the R code and its output from 10 and 11.
```{r}
length(seq_1_to_100)
length(seq_1_to_25)
sum(100, 25)
```
12. Statements can be grouped together using braces ‘{’ and ‘}’. A group of statements
is sometimes called a block. Single statements are evaluated when a new line is typed
at the end of the syntactically complete statement. Blocks are not evaluated until a
new line is entered after the closing brace.
Enter this statement:
x <- {0 + x + 5 + }

Describe the output.
- Error: unexpected '}' in "x <- {0 + x + 5 + }"

13. *Set up a vector named score, consisting of 72, 86, 92, 63, 88, 89, 91, 92, 75,
75 and 77. To access individual elements of an atomic vector, one generally uses the
x[i] construction.
Find x[2] and x[3]. Write the R code and its output.
```{r}
score <- c(72, 86, 92, 63, 88, 89, 91, 92, 75,
75, 77)
element_1 <- score[2]
element_2 <- score[3]
element_1
element_2
```
14. *Create a vector a = c(1,2,NA,4,NA,6,7).
a. Change the NA to 999 using the codes print(a,na.print="-999").

b. Write the R code and its output. Describe the output.
```{r}
a = c(1, 2, 999, 4, 999, 6, 7)
print(a, na.print="-999")
```
- the value of NA when it was not changed was -999. but when it was changed to 999, the value becomes 999 too.


15. A special type of function calls can appear on the left hand side of the assignment
operator as in > class(x) <- "foo".
Follow the codes below:
name = readline(prompt="Input your name: ")
age = readline(prompt="Input your age: ")
print(paste("My name is",name, "and I am",age ,"years old."))
print(R.version.string)
What is the output of the above code?
```{r}
name = readline(prompt="Input your name: ")
age = readline(prompt="Input your age: ")
print(paste("My name is",name, "and I am",age ,"years old."))
print(R.version.string)
```

