---
title: "RWorksheet_Camarista#4b"
author: "John Lyxton Camarista"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Using Loop Function
##for() loop
  1. 1. Using the for loop, create an R script that will display a 5x5 matrix as shown in Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
Hint Use abs() function to get the absolute value
```{r}
vectorA <- c(1, 2, 3, 4, 5)
matrixA <- matrix(0, nrow = 5, ncol = 5)

for(i in 1:5){

    matrixA[i, ] <- abs(vectorA - i)
}
matrixA
```
  2. Print the string "*" using for() function. The output should be the same as shown in Figure
```{r}
for(i in 1:5){

    cat(rep("*", i), "\n")
  
}
```
  3. Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r}
#There's an error during knitting in this particular chunk. So I just commented the prompt script an initialize 4 as the temporary user input.
#start <- as.integer(readline(prompt = "Enter the starting number: "))
start <- 4

a <- start
b <- start + 1

cat(a, "\n")

repeat {
  cat(b, "\n")
  
  next_term <- a + b
  
  if (next_term > 500) {
    break
  }
  
  a <- b
  b <- next_term
}

```


  
  
