---
title: "RWorksheet_Camarista#4a"
author: "John Lyxton Camarista"
date: "2024-10-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. The table below shows the data about shoe size and height. Create a data frame.
```{r}
height <- c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 67.0, 71.0, 
            71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 
            69.0, 70.0)

shoeSize <- c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5, 
              13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0)

shoeData <- data.frame(Shoe_Size = shoeSize, Height = height)

#got help from chat gpt to achieve the same layout from the Figure 1: Household Data, 
#having 2 halves of the data frame.
half <- ceiling(nrow(shoeData) / 2)
first_half <- shoeData[1:half, ]
second_half <- shoeData[(half+1):nrow(shoeData), ]

combined <- cbind(first_half, second_half)

combined
print(str(shoeData))
```

- a. Describe the data.
  - The data is consist of 28 rows and columns(Shoe Size and Height) with numerical values.   
  
  
- b. Create a subset by males and females with their corresponding shoe size and height.
What its result? Show the R scripts.
```{r}
gender <- c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", "M", "F", "M", "M", "M", 
            "M", "F", "F", "M", "F", "F", "M", "M", "F", "M", "M", "M", "M")

withGender <- cbind(shoeData, Gender = gender)

half <- ceiling(nrow(withGender) / 2)
first_half <- withGender[1:half, ]
second_half <- withGender[(half+1):nrow(withGender), ]

combinedWithGender <- cbind(first_half, second_half)

combinedWithGender
```
  - c. Find the mean of shoe size and height of the respondents. Write the R scripts and its
result.
```{r}
dataMeanShoeSize <- round(mean(withGender$Shoe_Size),2)
print(paste("The mean of the Shoe size is ", dataMeanShoeSize))

dataMeanHeight <- round(mean(withGender$Height),2)
print(paste("The mean of the Height is ", dataMeanHeight))
```
- d. Is there a relationship between shoe size and height? Why?
  - "Yes, there is. As height increases, shoe size also increases. This increase in shoe or foot size is necessary to support the person's relative height. Similarly, when constructing tall buildings, a deep foundation is required."
  
  
## 2. Construct character vector months to a factor with factor() and assign the result to factor_months_vector. Print out factor_months_vector and assert that R prints out the factor levels below the actual values.
  
  Consider data consisting of the names of months:
"March", "April", "January", "November", "January", "September", "October", "September", "November", "August", "January", "November", "November", "February", "May", "August", "July", "December", "August", "August", "September", "November", "February", "April"
  
```{r}
months <- c("March", "April", "January", "November", "January", "September", "October", 
            "September", "November", "August", "January", "November", "November", "February", 
            "May", "August", "July", "December", "August", "August", "September", "November", 
            "February", "April")

factor_months_vector <- factor(months)
factor_months_vector
```

## 3. Then check the summary() of the months_vector and factor_months_vector. Interpret the results of both vectors. Are they both equally useful in this case?
```{r}
summary_factor_months <- summary(factor_months_vector)
summary_factor_months
```

## 4. Create a vector and factor for the table below.
```{r}

direction <- c("East", "West", "North")
frequency <- c(1, 4, 3)

factor_data <- factor(direction, levels = c("East", "West", "North"))

direction_data <- data.frame(Direction = factor_data, Frequency = frequency)

direction_data

```

## 5. Enter the data below in Excel with file name = import_march.csv
-a. Import the excel file into the Environment Pane using read.table() function.
Write the code.
```{r}
import_march_data <- read.table("import_march.csv", header = TRUE, sep = ",")
```
-b. View the dataset. Write the R scripts and its result.
```{r}
import_march_data
```

# Using Conditional Statements (IF-ESE)
## 6. Fyll Search

-Exhaustive search is a methodology for finding an answer by exploring all possible cases. When trying to find a desired number in a set of given numbers, the method of finding the corresponding number by checking all elements in the set one by one can be called an exhaustive search. Implement an exhaustive search function that meets the input/output conditions below.
-a. Create an R Program that allows the User to randomly select numbers from 1 to 50. Then display the chosen number. If the number is beyond the range of the selected choice, it will have to display a string “The number selected is beyond the range of 1 to 50”. If number 20 is inputted by the User, it will have to display “TRUE”, otherwise display the input number.

```{r}
num_selected <- readline(prompt = "Enter a number between 1 to 50: ")
if(num_selected > 50){
  print("The number selected is beyond the range of 1 to 50")
}else if(num_selected == 20){
  print("TRUE")
}else{
  num_selected
}


```

  
  
  
  
  
  
  
  
  
  
  
  
  
  