---
title: "RWorksheet_Camarista#4a"
author: "John Lyxton Camarista"
date: "2024-10-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. The table below shows the data about shoe size and height. Create a data frame.
```{r}
height <- c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 67.0, 71.0, 
            71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 
            69.0, 70.0)

shoeSize <- c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5, 
              13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0)

shoeData <- data.frame(Shoe_Size = shoeSize, Height = height)

#got help from chat gpt to achieve the same layout from the Figure 1: Household Data, 
#having 2 halves of the data frame.
half <- ceiling(nrow(shoeData) / 2)
first_half <- shoeData[1:half, ]
second_half <- shoeData[(half+1):nrow(shoeData), ]

combined <- cbind(first_half, second_half)

combined
print(str(shoeData))
```

- a. Describe the data.
  - The data is consist of 28 rows and columns(Shoe Size and Height) with numerical values.  
  
  
- b. Create a subset by males and females with their corresponding shoe size and height.
What its result? Show the R scripts.
```{r}
gender <- c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", "M", "F", "M", "M", "M", 
            "M", "F", "F", "M", "F", "F", "M", "M", "F", "M", "M", "M", "M")

withGender <- cbind(shoeData, Gender = gender)

half <- ceiling(nrow(withGender) / 2)
first_half <- withGender[1:half, ]
second_half <- withGender[(half+1):nrow(withGender), ]

combinedWithGender <- cbind(first_half, second_half)

combinedWithGender
```
  - c. Find the mean of shoe size and height of the respondents. Write the R scripts and its
result.
```{r}
dataMeanShoeSize <- round(mean(withGender$Shoe_Size),2)
print(paste("The mean of the Shoe size is ", dataMeanShoeSize))

dataMeanHeight <- round(mean(withGender$Height),2)
print(paste("The mean of the Height is ", dataMeanHeight))
```
- d. Is there a relationship between shoe size and height? Why?
  - "Yes, there is. As height increases, shoe size also increases. This increase in shoe or foot size is necessary to support the person's relative height. Similarly, when constructing tall buildings, a deep foundation is required."
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  