---
title: "RWorksheet_Camarista#4b"
author: "John Lyxton Camarista"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Using Loop Function
##for() loop
  1. 1. Using the for loop, create an R script that will display a 5x5 matrix as shown in Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
  Hint Use abs() function to get the absolute value
```{r}
vectorA <- c(1, 2, 3, 4, 5)
matrixA <- matrix(0, nrow = 5, ncol = 5)

for(i in 1:5){

    matrixA[i, ] <- abs(vectorA - i)
}
matrixA
```
  2. Print the string "*" using for() function. The output should be the same as shown in Figure
```{r}
for(i in 1:5){

    cat(rep("*", i), "\n")
  
}
```
  3. Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r}
#There's an error during knitting in this particular chunk. So I just commented the prompt script an initialize 4 as the temporary user input.
#start <- as.integer(readline(prompt = "Enter the starting number: "))
start <- 4

a <- start
b <- start + 1

cat(a, "\n")

repeat {
  cat(b, "\n")
  
  next_term <- a + b
  
  if (next_term > 500) {
    break
  }
  
  a <- b
  b <- next_term
}

```

##Using Basic Graphics (plot(),barplot(),pie(),hist())
  4. Import the dataset as shown in Figure 1 you have created previously.
    - a. What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? Show your codes and its result
```{r}
ShoeSize_Data <- read.csv("ShoeSize.csv")
head(ShoeSize_Data)
```
    - b. Create a subset for gender(female and male). How many observations are there in Male? How about in Female? Write the R scripts and its output.
```{r}
male_sub <- subset(ShoeSize_Data, gender == "M", select = gender)
male_count <- nrow(male_sub)
print(paste("There are", male_count, "males"))

female_sub <- subset(ShoeSize_Data, gender == "F", select = gender)
female_count <- nrow(female_sub)
print(paste("There are", female_count, "females"))
```
    - c.Create a graph for the number of males and females for Household Data. Use plot(), chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and its result.
```{r}
gender_counts <- c(male_count, female_count)
gender_labels <- c("Male", "Female")
barplot(
  gender_counts,
  names.arg = gender_labels,
  main = "Number of Males and Females in Household Data",
  xlab = "Gender",
  ylab = "Count",
  col = c("blue", "#FF00FF")
)

```

##5. The monthly income of Dela Cruz family was spent on the following:
  - a. Create a piechart that will include labels in percentage.Add some colors and title of the chart. Write the R scripts and show its output.
  
  
```{r}
expenses <- c(Food = 60, Electricity = 10, Savings = 5, Miscellaneous = 25)

percent_labels <- paste(names(expenses), round(expenses / sum(expenses) * 100, 1), "%")

pie(
  expenses,
  labels = percent_labels,
  main = "Monthly Expenditure of Dela Cruz Family",
  col = c("lightblue", "lightgreen", "yellow", "pink")
)

```

##6. Use the iris dataset.
  - a. Check for the structure of the dataset using the str() function. Describe what you have seen in the output.
```{r}
# Load the iris dataset
data(iris)

# Check the structure of the dataset
str(iris)

```
  - b. Create an R object that will contain the mean of the sepal.length, sepal.width,petal.length,and petal.width. What is the R script and its result?
```{r}
irisMean <- colMeans(iris[, 1:4])
irisMean
```
  - c. Create a pie chart for the Species distribution. Add title, legends, and colors. Write the R script and its result.
```{r}
species_counts <- table(iris$Species)

pie(
  species_counts,
  labels = paste(names(species_counts), round(species_counts / sum(species_counts) * 100, 1), "%"),
  main = "Species Distribution in Iris Dataset",
  col = c("lightblue", "lightgreen", "lightpink")
)

legend("topright", legend = names(species_counts), fill = c("lightblue", "lightgreen", "lightpink"))

```
  - d. Subset the species into setosa, versicolor, and virginica. Write the R scripts and show the last six (6) rows of each species.
```{r}
setosa <- subset(iris, Species == "setosa")
versicolor <- subset(iris, Species == "versicolor")
virginica <- subset(iris, Species == "virginica")

tail(setosa)
tail(versicolor)
tail(virginica)

```
  - e. Create a scatterplot of the sepal.length and sepal.width using the different  species(setosa,versicolor,virginica). Add a title = “Iris Dataset”, subtitle = “Sepal width and length, labels for the x and y axis, the pch symbol and colors should be based on the species.
    Hint: Need to convert to factors the species to store categorical variables.
```{r}
plot(
  iris$Sepal.Length, iris$Sepal.Width,
  col = as.numeric(iris$Species),
  pch = as.numeric(iris$Species),
  main = "Iris Dataset",
  sub = "Sepal Width and Length",
  xlab = "Sepal Length",
  ylab = "Sepal Width"
)

legend("topright", legend = levels(iris$Species), col = 1:3, pch = 1:3)

```
  - f. Interpret the result
    *The Setosa species' sepal width has a relation to its sepal length. The longer the length, the wider it's width.
    *While the Versicolor and Virginica species tend to overlap in the middle. The longer their length, their width almost stays the same arounf 3.0.