---
title: "RWorksheet_Camarista#3b"
author: "John Lyxton Camarista"
date: "2024-10-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### 1. Create a data frame using the table below.
```{r}
# a. Write the codes.
dataFrame <- data.frame(
  Respondents = 1:20,
  Sex = c(2, 2, 1, 2, 2, 2, 2, 1, 2, 1, 2, 1, 2, 2, 1, 2, 1, 2, 1, 2),
  FathersOccupation = c(1, 3, 3, 3, 1, 2, 2, 3, 1, 3, 1, 3, 1, 2, 2, 3, 1, 3, 2, 1),
  PersonsAtHome = c(5, 7, 3, 8, 9, 6, 7, 8, 4, 6, 8, 5, 6, 9, 1, 6, 11, 7, 7, 6),
  SiblingsAtSchool = c(6, 4, 3, 1, 8, 5, 5, 2, 2, 3, 1, 2, 2, 1, 1, 6, 5, 1, 2, 2),
  TypesOfHouses = c(1, 3, 1, 3, 3, 2, 1, 3, 3, 2, 3, 2, 1, 3, 2, 3, 1, 3, 3, 2))
dataFrame
```
```{r}
#b. Describe the data. Get the structure or the summary of the data
str(dataFrame)  
summary(dataFrame)
```
  * the data frame has 20 rows and 6 columns.
    + There are 20 respondents
    + In sex, 1 is for male and 2 is for female
    + In father's occupation, 1 is for farmer, 2 for driver, and 3 for others
    + The Person's at home shows how many people living in the respondent's house
    + Siblings at schools shows how many siblings of the respondents are still attending school
    + In types of house, 1 is for wood, 2 is for semi-concrete ans 3 for concrete  
    
```{r}
#c. Is the mean number of siblings attending is 5?
meanSiblings <- mean(dataFrame$SiblingsAtSchool)

print(paste(meanSiblings, "is the mean of Sibling attending school"))
```
```{r}
#d. Extract the 1st two rows and then all the columns using the subsetting functions.
#Write the codes and its output.
subset(dataFrame[1:2, ])
```
```{r}
#e. Extract 3rd and 5th row with 2nd and 4th column. Write the codes and its result.
dataFrame[c(3, 5), c(2, 4)]
```
```{r}
#f. Select the variable types of houses then store the vector that results as types_houses.
#Write the codes.
types_houses <- dataFrame$TypesOfHouses
types_houses
```
```{r}
#g. Select only all Males respondent that their father occupation was farmer. 
#Write the codes and its output.
male_farmer <- dataFrame[dataFrame$Sex == 1 & dataFrame$FathersOccupation == 1, ]
male_farmer
```
```{r}
#h. Select only all females respondent that have greater than or equal to 5 number of siblings attending school. 
#Write the codes and its outputs.
female_siblings <- dataFrame[dataFrame$Sex == 2 & dataFrame$SiblingsAtSchool >= 5, ]
female_siblings
```
  #### 2. Write a R program to create an empty data frame. Using the following codes:
```{r}
df = data.frame(Ints=integer(),

Doubles=double(), Characters=character(),
Logicals=logical(),
Factors=factor(),
stringsAsFactors=FALSE)

print("Structure of the empty dataframe:")
print(str(df))
```
  * The structure of the empty data frame shows no data but has predefined columns with their respective data types.  
  
#### 3. Create a .csv file of this. Save it as HouseholdData.csv
  
```{r}
#a. Import the csv file into the R environment. Write the codes.
HouseholdData <- read.csv("HouseholdData.csv")
HouseholdData
```
```{r}
#b. Convert the Sex into factor using factor() function and change it into integer.
#[Legend: Male = 1 and Female = 2]. Write the R codes and its output.
HouseholdData$Sex <- as.numeric(factor(HouseholdData$Sex, levels = c("Male", "Female")))
HouseholdData
```
```{r}
#c. Convert the Type of Houses into factor and change it into integer. 
#[Legend: Wood = 1; Concrete = 2; Semi-Concrete = 3]. 
#Write the R codes and its output.
HouseholdData$Types_of_Houses <- as.numeric(factor(HouseholdData$Types_of_Houses, levels = c("Wood", "Congrete", "Semi-Congrete")))
HouseholdData
```
```{r}
#d. On father’s occupation, factor it as Farmer = 1; Driver = 2; and Others = 3. 
#What is the R code and its output?
HouseholdData$Fathers_Occupation <- as.character(factor(HouseholdData$Fathers_Occupation, levels = c(1, 2, 3), labels = c("Farmer", "Driver", "Others")))

HouseholdData
```
```{r}
#e. Select only all females respondent that has a father whose occupation is driver. 
#Write the codes and its output.
subset(HouseholdData[,c(2:3)], Sex == 2 & Fathers_Occupation == "Driver")

```
```{r}
#f. Select the respondents that have greater than or equal to 5 number of siblings attending school. 
#Write the codes and its output.
subset(HouseholdData[,c(1,5)], Siblings_at_School >= 5)
```
#### 4. Interpret the Graph
  - The graph is a Bar graph titled: "Sentiments of Tweets per day". 
    + The graph date ranges from July 14, 2020 to July 22, 2020.
    + Each day has 3 bars: *Red* for Negative, *Yellow* for Neutral, and *Blue* for Positive.
    + The peak of red bar was in July 15. 
    + The peak of yellow bar was also in July 15.
    + And the peak of blue bar was in July 21.
    + The Y-axis of the graph represents the number of tweets. Ranging from 0 to more than 3500.
    + The highest number of Tweets was on July 15, 2020. With around 4000 **Negative** tweets.
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
    

