---
title: "RWorksheet_Camarista#3b"
author: "John Lyxton Camarista"
date: "2024-10-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### 1. Create a data frame using the table below.
```{r}
# a. Write the codes.
dataFrame <- data.frame(
  Respondents = 1:20,
  Sex = c(2, 2, 1, 2, 2, 2, 2, 1, 2, 1, 2, 1, 2, 2, 1, 2, 1, 2, 1, 2),
  FathersOccupation = c(1, 3, 3, 3, 1, 2, 2, 3, 1, 3, 1, 3, 1, 2, 2, 3, 1, 3, 2, 1),
  PersonsAtHome = c(5, 7, 3, 8, 9, 6, 7, 8, 4, 6, 8, 5, 6, 9, 1, 6, 11, 7, 7, 6),
  SiblingsAtSchool = c(6, 4, 3, 1, 8, 5, 5, 2, 2, 3, 1, 2, 2, 1, 1, 6, 5, 1, 2, 2),
  TypesOfHouses = c(1, 3, 1, 3, 3, 2, 1, 3, 3, 2, 3, 2, 1, 3, 2, 3, 1, 3, 3, 2))
dataFrame
```
```{r}
#b. Describe the data. Get the structure or the summary of the data
str(dataFrame)  
summary(dataFrame)
```
  * the data frame has 20 rows and 6 columns.
    + There are 20 respondents
    + In sex, 1 is for male and 2 is for female
    + In father's occupation, 1 is for farmer, 2 for driver, and 3 for others
    + The Person's at home shows how many people living in the respondent's house
    + Siblings at schools shows how many siblings of the respondents are still attending school
    + In types of house, 1 is for wood, 2 is for semi-concrete ans 3 for concrete  
    
```{r}
#c. Is the mean number of siblings attending is 5?
meanSiblings <- mean(dataFrame$SiblingsAtSchool)

print(paste(meanSiblings, "is the mean of Sibling attending school"))
```
```{r}
#d. Extract the 1st two rows and then all the columns using the subsetting functions.
#Write the codes and its output.
subset(dataFrame[1:2, ])
```
```{r}
#e. Extract 3rd and 5th row with 2nd and 4th column. Write the codes and its result.
dataFrame[c(3, 5), c(2, 4)]
```
```{r}
#f. Select the variable types of houses then store the vector that results as types_houses.
#Write the codes.
types_houses <- dataFrame$TypesOfHouses
types_houses
```
```{r}
#g. Select only all Males respondent that their father occupation was farmer. 
#Write the codes and its output.
male_farmer <- dataFrame[dataFrame$Sex == 1 & dataFrame$FathersOccupation == 1, ]
male_farmer
```
```{r}
#h. Select only all females respondent that have greater than or equal to 5 number of siblings attending school. 
#Write the codes and its outputs.
female_siblings <- dataFrame[dataFrame$Sex == 2 & dataFrame$SiblingsAtSchool >= 5, ]
female_siblings
```
  #### 2. Write a R program to create an empty data frame. Using the following codes:
```{r}
df = data.frame(Ints=integer(),

Doubles=double(), Characters=character(),
Logicals=logical(),
Factors=factor(),
stringsAsFactors=FALSE)

print("Structure of the empty dataframe:")
print(str(df))
```
  * The structure of the empty data frame shows no data but has predefined columns with their respective data types.  
  
#### 3. Create a .csv file of this. Save it as HouseholdData.csv
  
```{r}
#a. Import the csv file into the R environment. Write the codes.
HouseholdData <- read.csv("E:/Github/Data Science Worksheets/DataScience_Worksheets_Camarista/Worksheet#3/HouseholdData.csv")
HouseholdData
```
```{r}
#b. Convert the Sex into factor using factor() function and change it into integer.
#[Legend: Male = 1 and Female = 2]. Write the R codes and its output.
HouseholdData$sex <- factor(HouseholdData$sex, levels = c(1, 2), labels = c("Male", "Female"))
HouseholdData
```
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
    

